Investigating Fault-Proneness Through Hotfix Commits in the Light of Different Programming Languages

### 2016-02-20 msr_paper/main.tex
Fixed the number of programming languages in the abstract from 6 to 8
Fixed some typos and turning all headers into lower case

\begin{abstract}
Choosing the most suitable programming language is a challenging task for every project manager and developer. Often the success of a software project is determined by the stability of the system after a release. We analyse the prevalence of hotfix commits in various development areas as an indicator of which languages tend to produce the least amount of problems after software has been published. Based on the Boa September 2013 SourceForge dataset, we analysed the metadata from the 16 most popular project types and estimated the amount of deployed hotfixes to compare the fault-proneness of programming languages that are use across project types. Our results only identified 6 languages that are widely used across different domains, Java, C, C++, C\#, Perl, and PHP. Java has the least number of hotfixes per version update (0.18), while C\# required most hotfixes per version update (0.34). We identified the Security and Internet \& Communications project types as being the most reliable (i.e.\ requiring least amount of hotfixes per version update). The analysis did not reveal conclusive patterns for language classes (strongly vs.\ weakly typed and dynamic vs.\ static).
\end{abstract}

@@

\subsection{Identifying hotfix commits}
%The first major problem we had to solve was identifying actual hotfix commits. Due to Boa's limitations, we were not able to find commit history from hotfix branch. We were only capable of collecting information with regard to commits performed on the main branch. Therefore, we reviewed various papers to identify the most appropriate method of determining the type of the commit to extract only the ones related to a bug correction.\par

@@

\subsection{Research Questions}
\noindent Our study aimed to answer the following research questions:

\noindent\textbf{RQ1 (Language Popularity):} What languages are the most widely used across the top 16 most popular categories?

\noindent\textbf{RQ2 (Hotfixes vs Project Categories):} What project categories achieve the highest/lowest number of hotfixes per version update?

\noindent\textbf{RQ3 (Hotfixes vs Languages):} What languages achieve the highest/lowest number of hotfixes per version update?

\noindent\textbf{RQ4 (Hotfixes vs Language Types):} Are there any relationship between the occurrences of hotfixes and the type of the programming language?

@@

\section{Results}
We performed an empirical study on the SourceForge13 dataset aiming to find the answers for our 4 research questions. The findings of each research question is explained below.
\subsection{RQ1: Language Popularity}
Our investigation on the most widely used languages across the top 16 most popular categories revealed that there are surprisingly few such languages. We identified only 8 widely used languages: Java, C, C++, Perl, Python, PHP, C\#, and JavaScript. The criteria of choosing the languages is based on whether the language presents in most of the categories. We found that there is a significant drop in the amount of occurrence after JavaScript.  %One would expect JavaScript to appear in the list. However, it is present only in 7 categories and hence is ignored from the list. This may be due to the fact that JavaScript is mainly used only for website development while our 16 popular categories span over various aspects of software.

@@

\subsection{RQ2: Hotfixes vs Project Categories}
Next, we investigate which categories achieve the highest or lowest numbers of hotfixes per version update. We proceed to analyse why certain categories might have low numbers of hotfixes and vice versa.

@@

\subsection{RQ3: Hotfixes vs Languages}
Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has appeared to have the least average ratio of hotfixes per version update (0.176)} followed by JavaScript (0.209) and Perl (0.218). However, all the mentioned languages show marginally better results with 8\% difference comparing to other common languages taken under analysis: PHP(0.282), C++(0.286), C (0.305) and C\#(0.339). \textbf{C\# has the highest average ratio of hotfixes among all the languages.}

\subsection{RQ4: Hotfixes vs Language Types}
The interesting research question is comparing the type of each programming language by their average ratio of hotfixes. The classification in Table \ref{tab:1} is inspired by \cite{Ray2014}, where languages were compared based on the number of bug fixes in each language and were divided into three main Classes. The first class (Programming Paradigm) has the following categories: Procedural, Functional, Scripting. The second class is Compilation containing Static and Dynamic type. The final one, Type Class, separates languages into Strongly-typed and Weakly-typed. According to \cite{Ray2014}, strongly-typed languages experienced fewer bug fixes than its counterpart, weakly-typed. Static programming languages required fewer corrective commits than dynamic ones. And finally, functional languages encountered fewer bug fixes than procedural ones.

@@

\subsubsection*{Programming Paradigm: Procedural vs Scripting}
From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among Procedural (Java, C, C\#, C++) and Scripting (Perl, Python, PHP) across all the categories. This is possibly because the number of data points (languages) was too low to observe such behaviours and make any firm conclusions. %It was always the case that at least one of the languages would not follow a pattern. For some categories, the comparison would consider only one language from a certain category.

\subsubsection*{Compilation: Static vs Dynamic}
In this class, Static (Java, C, C\#, C++) and Dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear different of hotfix ratio observed between the two sets as can be seen from Table \ref{tab:1}. The reason was because the languages in these two categories coincided with the languages in the first two classes (Procedural and Scripting).

\subsubsection*{Type: Strongly-type vs Weakly-typed}
The third class compares strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages. Similarly to the previous two classes, we encountered no clear pattern across the categories. Furthermore, we found the same problem of having comparisons against only one member of a language class. Solely in the Communications category, we were able to detect some pattern, where strongly-typed languages performed slightly better. However, there were only two languages labelled as weak. Therefore, we cannot reach any satisfying conclusions based on this finding.

>>>>

\begin{abstract}
Choosing the most suitable programming language is a challenging task for every project manager and developer. Often the success of a software project is determined by the stability of the system after a release. We analyse the prevalence of hotfix commits in various development areas as an indicator of which languages tend to produce the least amount of problems after software has been published. Based on the Boa September 2013 SourceForge dataset, we analysed the metadata from the 16 most popular project types and estimated the amount of deployed hotfixes to compare the fault-proneness of programming languages that are use across project types. Our results only identified 8 languages that are widely used across different domains, Java, C, C++, C\#, Perl, Python, PHP, and JavaScript. Java has the least number of hotfixes per version update (0.18), while C\# required most hotfixes per version update (0.34). We identified the Security and Internet \& Communications project types as being the most reliable (i.e.\ requiring least amount of hotfixes per version update). The analysis did not reveal conclusive patterns for language classes (strongly vs.\ weakly typed and dynamic vs.\ static).
\end{abstract}

@@

\subsection{Identifying bug fix commits}
%The first major problem we had to solve was identifying actual hotfix commits. Due to Boa's limitations, we were not able to find commit history from hotfix branch. We were only capable of collecting information with regard to commits performed on the main branch. Therefore, we reviewed various papers to identify the most appropriate method of determining the type of the commit to extract only the ones related to a bug correction.\par

@@

\subsection{Approach for identifying hotfix commits}

\subsection{Research questions}
\noindent Our study aimed to answer the following research questions:

\noindent\textbf{RQ1 (language popularity):} What languages are the most widely used across the top 16 most popular categories?

\noindent\textbf{RQ2 (hotfixes vs project categories):} What project categories achieve the highest/lowest number of hotfixes per version update?

\noindent\textbf{RQ3 (hotfixes vs languages):} What languages achieve the highest/lowest number of hotfixes per version update?

\noindent\textbf{RQ4 (hotfixes vs language types):} Are there any relationship between the occurrences of hotfixes and the type of the programming language?

@@

\section{Results}
We performed an empirical study on the SourceForge13 dataset aiming to find the answers for our 4 research questions. The findings of each research question is explained below.
\subsection{RQ1: language popularity}
Our investigation on the most widely used languages across the top 16 most popular categories revealed that there are surprisingly few such languages. We identified only 8 widely used languages: Java, C, C++, Perl, Python, PHP, C\#, and JavaScript. The criteria of choosing the languages is based on whether the language presents in most of the categories. We found that there is a significant drop in the amount of occurrence after JavaScript.  %One would expect JavaScript to appear in the list. However, it is present only in 7 categories and hence is ignored from the list. This may be due to the fact that JavaScript is mainly used only for website development while our 16 popular categories span over various aspects of software.

@@

\subsection{RQ2: hotfixes vs project categories}
Next, we investigate which categories achieve the highest or lowest numbers of hotfixes per version update. We proceed to analyse why certain categories might have low numbers of hotfixes and vice versa.

@@

\subsection{RQ3: hotfixes vs languages}
Language comparison based on the average hotfix ratio over all the categories in Table 1 has shown that \textbf{Java has appeared to have the least average ratio of hotfixes per version update (0.176)} followed by JavaScript (0.209) and Perl (0.218). However, all the mentioned languages show marginally better results with 8\% difference comparing to other common languages taken under analysis: PHP(0.282), C++(0.286), C (0.305) and C\#(0.339). \textbf{C\# has the highest average ratio of hotfixes among all the languages.}

\subsection{RQ4: hotfixes vs language types}
Another interesting research question is comparing the type of each programming language by their average ratio of hotfixes. The classification in Table \ref{tab:1} is inspired by \cite{Ray2014}, where languages were compared based on the number of bug fixes in each language and were divided into three main Classes. The first class (Programming Paradigm) has the following categories: Procedural, Functional, Scripting. The second class is Compilation containing Static and Dynamic type. The final one, Type Class, separates languages into Strongly-typed and Weakly-typed. According to \cite{Ray2014}, strongly-typed languages experienced fewer bug fixes than its counterpart, weakly-typed. Static programming languages required fewer corrective commits than dynamic ones. And finally, functional languages encountered fewer bug fixes than procedural ones.

@@

\subsubsection*{Programming paradigm: procedural vs scripting}
From Table \ref{tab:1}, there is no clear trend or pattern of the hotfix ratio among Procedural (Java, C, C\#, C++) and Scripting (Perl, Python, PHP) across all the categories. This is possibly because the number of data points (languages) was too low to observe such behaviours and make any firm conclusions. %It was always the case that at least one of the languages would not follow a pattern. For some categories, the comparison would consider only one language from a certain category.

\subsubsection*{Compilation: static vs dynamic}
In this class, static (Java, C, C\#, C++) and dynamic (Perl, Python, PHP) languages led to the same conclusions of no clear different of hotfix ratio observed between the two sets as can be seen from Table \ref{tab:1}. The reason was because the languages in these two categories coincided with the languages in the first two classes (procedural and scripting).

\subsubsection*{Type: strongly-type vs weakly-typed}
The third class compares strongly-typed (Java, C\#, Perl, Python) and weakly-typed (C, C++, PHP) languages. Similarly to the previous two classes, we encountered no clear pattern across the categories. Furthermore, we found the same problem of having comparisons against only one member of a language class. Solely in the Communications category, we were able to detect some pattern, where strongly-typed languages performed slightly better. However, there were only two languages labelled as weak. Therefore, we cannot reach any satisfying conclusions based on this finding.
